---
title: "R Notebook"
output: html_notebook
---

```{r}
# libraries
library(tidyverse)
```

```{r}
# loading the dataset
raw = read_csv(
  here::here(
    "Datasets",
    "Laptop_price_estimator",
    "laptop_price.csv"
  )
)
```

```{r}
head(raw)
```

# **Data Cleaning**

I decided to start with the data cleaning so that we can do a proper Exploratory.

```{r}
# raw_df = raw_df[-1] -  dropping the laptop_id feature and commenting it out since we only need to do this once.

# Renaming the inches column to screen size
raw = raw %>% rename(screen_size = Inches)
# let's rename this to product type
raw = raw %>% rename(product_type = TypeName)
# column name formatting
raw = janitor::clean_names(raw)
```

The screen_resolution columns needs to be separated into 3 columns: resolution, display, and touchscreen

```{r}
# display
raw$display = gsub("\\d+x\\d+", "", raw$screen_resolution)

# creating the touchscreen column from display column
raw$touch_screen = ifelse(grepl("Touchscreen", raw$display), 1, 0)
raw$touch_screen = factor(raw$touch_screen, levels = c(0, 1), labels = c("No", "Yes"))

# display - continued
raw$display = gsub("Touchscreen", "",raw$display)
raw$display = gsub("/", "",raw$display)
raw$display = trimws(raw$display)
raw$display = ifelse(raw$display=="IPS Panel   4K Ultra HD", "IPS Panel 4K Ultra HD", raw$display)
raw$display = ifelse(raw$display=="", "None",raw$display)
raw$display = gsub("Touchscreen", "",raw$display)
raw$display = gsub("/", "",raw$display)
raw$display = trimws(raw$display)
raw$display = ifelse(raw$display=="IPS Panel   4K Ultra HD", "IPS Panel 4K Ultra HD", raw$display)
raw$display = ifelse(raw$display=="", "None",raw$display)  
```

```{r}
# resolution
raw$resolution = str_extract(raw$screen_resolution, "\\d+x\\d+")
```

There are three major CPU brands in our dataset. It would be beneficial to simplify use the brand names.

```{r}
# The CPU column
split_cpu = str_split(raw$cpu, " ")
# spliting the feature into  a list separate arguments
split_cpu = str_split(raw$cpu, " ")
# creating the CPU brand column
raw$cpu = as.character(lapply(split_cpu, function(x) paste(x[1])))
raw$processor_speed = map(split_cpu, tail, n = 1)
raw$processor_speed = gsub("GHz", "", raw$processor_speed)
raw$processor_speed = as.numeric(as.character(raw$processor_speed))
```

```{r}
# The Ram column 
raw$ram = gsub("GB", "", raw$ram)

raw$ram = as.numeric(as.character(raw$ram))
```

```{r}
# The memory column
# spliting the column into  a list separate arguments
split_memory = str_split(raw$memory, " ")
# creating new list with the desired arguments
drive_capacity = lapply(split_memory, function(x) paste(x[1], collapse = ""))
drive_type = lapply(split_memory, function(x) paste(x[2],x[3], collapse = ""))
extra_drive_capacity = lapply(split_memory, function(x) paste(x[3],x[4],x[5], collapse = ""))
# transforming the list into dataframes
drive_capacity = as.data.frame(drive_capacity)
drive_type = as.data.frame(drive_type)
extra_drive_capacity = as.data.frame(extra_drive_capacity)
```

The memory column was split into the hard drive capacity column and hard drive type.

```{r}
# drive capacity column
drive_capacity = pivot_longer(drive_capacity,
                      cols = colnames(drive_capacity),
                      names_to = "var",
                      values_to = "capacity")

drive_capacity$capacity = gsub("GB", "", drive_capacity$capacity)

drive_capacity$capacity = gsub("TB", "000", drive_capacity$capacity)

drive_capacity$capacity = ifelse(drive_capacity$capacity=="1TB"|
                                drive_capacity$capacity=="1.0000", 
                                1000,drive_capacity$capacity)

raw$drive_capacity = drive_capacity$capacity

raw$drive_capacity = as.numeric(as.character(raw$drive_capacity))
```

```{r}
# drive type column
drive_type = as.data.frame(drive_type)
drive_type = pivot_longer(drive_type,
                          cols = colnames(drive_type),
                          names_to = "var",
                          values_to = "type")
drive_type$type = gsub("NA", "", drive_type$type)
drive_type$type = gsub("\\+", "", drive_type$type)
drive_type$type = trimws(drive_type$type)

raw$drive_type = drive_type$type
```

```{r}
# extra drive capacity
extra_drive_capacity = as.data.frame(extra_drive_capacity)

extra_drive_capacity = pivot_longer(extra_drive_capacity,
                          cols = colnames(extra_drive_capacity),
                          names_to = "var",
                          values_to = "extra")


extra_drive_capacity$extra = ifelse(grepl("NA", extra_drive_capacity$extra), 0, 1)


raw$extra_drive_capacity = factor(extra_drive_capacity$extra, 
                                 levels = c(0,1),
                                 labels = c("No","Yes"))
```

GPU - There are three major GPU brands in our dataset. It would be beneficial to simplify this column as well.

```{r}
# GPU
split_gpu = str_split(raw$gpu, " ")
gpu = lapply(split_gpu, function(x) paste(x[1], collapse = ""))
gpu = as.data.frame(gpu)
gpu = gpu %>% 
  pivot_longer(cols = colnames(gpu),
               names_to = "var",
               values_to = "brand")

raw$gpu = gpu$brand
```

```{r}
# weight 
raw$weight = gsub("kg", "", raw$weight)
raw$weight = as.numeric(as.character(raw$weight))
```

```{r}
# product
raw$product <- iconv(raw$product, "UTF-8", "ASCII", sub="")
raw$product <-  sub("\\(.*", "",raw$product)
```

# **EDA**

```{r}
# subsetting the desired columns
df = raw %>% select(
  company, 
  product, 
  product_type,
  screen_size,
  resolution,
  display,
  touch_screen,
  cpu,
  processor_speed,
  ram,
  drive_capacity,
  drive_type,
  extra_drive_capacity,
  gpu,
  op_sys,
  weight,
  price_euros
  )
```

```{r}
skimr::skim(df)
```

No missing values were identified. However, many features need to be formatted properly. So, let\'s start with the numerical variables.

### Numerical variables

-   Price

```{r}
# visualization
df %>% ggplot(aes(price_euros) ) +
  geom_histogram(fill="steelblue", binwidth = 500) +
  labs(title = "distribution of the prices",
       x = "prices",
       y = "count of values")
```

-   Screen size

```{r}
# visualization
df %>% ggplot(aes(screen_size) ) +
  geom_histogram(fill="steelblue", binwidth = 2 ) +
  labs(title = "distribution of the computer screen sizes",
       x = "screen size",
       y = "count of values")
```

-   CPU Processor Speed

```{r}
df %>% 
  ggplot(aes(processor_speed)) +
  geom_histogram(fill="steelblue")
```

-   Weight

```{r}
df %>% 
  ggplot(aes(weight)) +
  geom_histogram(fill="steelblue")
```

-   Drive capacity

```{r}
df %>% 
  ggplot(aes(drive_capacity)) +
  geom_histogram(fill="steelblue")
```

### Categorical variables

First, let's create a function to facilitate the visualization of the categorical variables

```{r}
# function to visualize each unique categorical variables
create_bar_chart <- function(data, var) {
  data %>%
    group_by(!!sym(var)) %>%
    summarise(count = n()) %>%
    ggplot(aes(fct_reorder(!!sym(var), count, .desc=TRUE), count)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(x = var, 
         y = "Count",
         title = paste0("Count by ", var))
  }
```

```{r}
# function to visualize the average price across each categorical variables
graph_price_by <- function(data, var) {
  data %>%
    group_by(!!sym(var)) %>%
    summarise(avg = mean(price_euros)) %>%
    ggplot(aes(fct_reorder(!!sym(var), avg, .desc=TRUE), avg)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(x = var, 
         y = "Average Price",
         title = paste0("Average Price by ", var))
}
```

-   Company

```{r}
create_bar_chart(df, "company") +
  coord_flip()
```

```{r}
graph_price_by(df, "company") +
  coord_flip()
```

-   Product type

```{r}
create_bar_chart(df, "product_type")
```

```{r}
graph_price_by(df, "product_type")
```

-   Resolution

```{r}
create_bar_chart(df, "resolution") +
  coord_flip()
```

```{r}
graph_price_by(df, "resolution") +
  coord_flip()
```

-   Display

```{r}
create_bar_chart(df, "display") +
  coord_flip() 
```

```{r}
graph_price_by(df, "display") +
  coord_flip()
```

-   CPU

```{r}
create_bar_chart(df, "cpu")
```

```{r}
graph_price_by(df, "cpu")
```

-   Drive types

```{r}
create_bar_chart(df, "drive_type")
```

```{r}
graph_price_by(df, "drive_type")
```

-   GPU

```{r}
create_bar_chart(df, "gpu")
```

```{r}
graph_price_by(df, "gpu")
```

-   Operating system

```{r}
create_bar_chart(df, "op_sys")
```

```{r}
graph_price_by(df, "op_sys")
```

-   Product

```{r}
# top 10 Products
df %>%
  count(product, sort = TRUE) %>% 
  head(10) %>% 
  ggplot(aes(fct_reorder(product, n), n)) +
  geom_col(fill="steelblue") +
  coord_flip() +
  labs(
      title = "Top 10 Products",
      x = "Product",
      y = "Count"
    )
```

```{r}
# Avg price by products
df %>%
  group_by(product, sort = TRUE) %>% 
  summarise(avg = mean(price_euros)) %>% 
  head(10) %>% 
  ggplot(aes(fct_reorder(product, avg), avg)) +
  geom_col(fill="steelblue") +
  coord_flip() +
  labs(
      title = "Avg price for 10 Products",
      x = "Product",
      y = "Avg. Price"
    )
```

-   Touch screen

```{r}
df %>% 
  ggplot(aes(touch_screen, fill= touch_screen)) +
  geom_bar() +
    labs(
      title = "Count of Products with Touch screen",
      x = "Touch Screen",
      y = "Ccount"
    )
```

```{r}
df %>% 
  group_by(touch_screen) %>% 
  summarise(avg = mean(price_euros)) %>% 
  ggplot(aes(touch_screen, avg)) +
  geom_col(fill = "steelblue") +
    labs(
      title = "Avg price for Products with Touch Screen",
      x = "Touch Screen",
      y = "Avg. Price"
    )
```

-   Extra drive capacity

```{r}
df %>% 
  ggplot(aes(extra_drive_capacity, fill= extra_drive_capacity)) +
  geom_bar() +
   labs(
      title = "Count of Products with Extra Drive Capacity",
      x = "Touch Screen",
      y = "Ccount"
    )
```

```{r}
df %>% 
  group_by(extra_drive_capacity) %>% 
  summarise(avg = mean(price_euros)) %>% 
  ggplot(aes(extra_drive_capacity, avg)) +
  geom_col(fill = "steelblue") +
    labs(
      title = "Avg price Product has Extra Drive Capacity",
      x = "Extra Drive Capacity",
      y = "Avg. Price"
    )
```
